\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}

\hypersetup{
  pdftitle={Detecting Strategically Decomposed Malicious Code Changes with Large Language Model Reviewers},
  pdfauthor={Anonymous pre-outcome manuscript},
  pdflang={en-AU},
  pdfdisplaydoctitle=true
}

\newcommand{\study}{LLM Review Sequences}
\newcommand{\code}[1]{\texttt{#1}}

\title{Detecting Strategically Decomposed Malicious Code Changes\\
with Large Language Model Reviewers}
\author{Anonymous pre-outcome manuscript}
\date{17 July 2026}

\begin{document}
\maketitle

\begin{abstract}
Large language models are increasingly used to review code changes, yet most
evaluations present a vulnerable function, file, or repository as one unit. A
malicious change can instead be decomposed across submissions that appear
individually routine but compose into an unsafe final state. We describe a
prospectively specified, controlled study of whether this decomposition reduces detection
and whether detection differs between pull-request gates and trunk workflows
that treat main as untrusted until pipeline review completes. The study uses 400
matched malicious-benign scenario pairs generated from 200 structural templates
in eight abstract policy families. Every program is synthetic, executes only in
memory, and has no network, file-system, credential, persistence, deployment, or
destructive capability. Atomic and three-submission variants have identical
final trees and changed-line totals. A factorial design varies intent,
decomposition, workflow framing, and local versus cumulative context across
three fixed model routes and three repeated trials. Primary outcomes are timely
blocking and paired risk differences; secondary outcomes include false-positive
rate, precision, calibration, localization, time to detection, and consistency.
Family-stratified structural-template bootstrapping, missing-data bounds, and
complete-pair sensitivity analyses are frozen before ground truth is joined to
responses. This pre-outcome manuscript reports the design and analysis plan; it
contains no confirmatory model outcomes.
\end{abstract}

\section{Introduction}

Automated code review is usually evaluated as a classification problem: a model
receives a code artifact and predicts whether it contains a defect. Real delivery
systems expose a different unit of work. Reviewers see submissions over time,
with evidence constrained by pull-request boundaries, commit history, repository
state, and pipeline policy. An unsafe capability may emerge only when several
changes compose. Each earlier change can be necessary for the final failure
without being unsafe in isolation.

This distinction matters for both pull-request and trunk-based delivery. A
pull-request gate asks whether a proposed change may enter the protected branch.
A trunk workflow can permit a commit onto main while still treating main as
untrusted until deterministic and model-based pipeline gates complete. Those
workflows differ in policy, but empirical comparisons often also change the
evidence shown to the reviewer. A useful experiment must therefore vary workflow
framing independently from information availability.

We study whether an LLM reviewer recognizes a composition before its activation
boundary. The benchmark represents harm as an abstract policy assertion: an
unprivileged synthetic actor receives a protected synthetic capability. It does
not implement an exploit or interact with a real target. The design makes three
contributions:

\begin{enumerate}[leftmargin=*]
  \item a causally validated benchmark in which every malicious split step is
  necessary, while atomic and split variants reach the same final tree;
  \item a factorial comparison of decomposition, workflow policy, and reviewer
  context that separates workflow effects from information loss; and
  \item a frozen sequence-level analysis with paired benign controls,
  template-respecting uncertainty, missingness bounds, and reproducible ledgers.
\end{enumerate}

\section{Related Work}

Yu et al. evaluated seven LLMs and multiple prompt forms for security code
review, comparing them with static analysis and examining factors associated
with detection quality \cite{yu2024securityreview}. Their results motivate fixed
prompts, false-positive reporting, and explicit context controls. Our study
changes the unit of analysis from one review artifact to a sequence of
submissions.

VulDetectBench separates vulnerability identification, classification, and
localization, showing that aggregate accuracy can conceal weaknesses in detailed
analysis \cite{liu2024vuldetectbench}. We likewise distinguish timely blocking,
calibration, severity, and localization. Zhou et al. compare repository-level
LLMs and static analysis tools and report a recall-false-positive tradeoff
\cite{zhou2024sast}. We retain a deterministic baseline where an installed tool
applies, but do not claim that syntactic static analysis and semantic review are
equivalent tasks.

Field evidence also indicates that review context and interaction design affect
developer trust and usefulness. Adalsteinsson et al. report context switching,
missing context, false positives, and trust as practical concerns in an
industrial study of LLM-assisted review \cite{adalsteinsson2025workflows}. Basic
and Giaretta synthesize evidence on LLM code-security benefits, false findings,
prompting, and failure modes \cite{basic2024review}. Our controlled study
complements that work by isolating temporal composition and workflow framing.

\section{Research Questions and Hypotheses}

The primary endpoint is \emph{blocked by activation}: at least one review at or
before the prespecified activation boundary returns \code{block}. The research
questions are:

\begin{description}[leftmargin=3.2em,style=nextline]
  \item[RQ1] Can LLM reviewers discriminate malicious sequences from matched
  benign sequences?
  \item[RQ2] Does strategic decomposition across three submissions reduce timely
  detection relative to an atomic submission with the same final tree?
  \item[RQ3] Does detection differ between pull-request and untrusted-main trunk
  workflow framing, and does workflow moderate the decomposition effect?
  \item[RQ4] Does cumulative submission history mitigate any decomposition
  penalty relative to local current-submission evidence?
\end{description}

The prespecified hypotheses are directional except where noted:

\begin{description}[leftmargin=3.2em,style=nextline]
  \item[H1] The malicious block rate exceeds the benign block rate.
  \item[H2] Under local context, the malicious block rate is lower for split
  sequences than for atomic submissions.
  \item[H3a] The marginal trunk-minus-pull-request risk difference is equivalent
  within -0.10 and 0.10.
  \item[H3b] The decomposition-by-workflow interaction differs from zero.
  \item[H4] Cumulative context attenuates the split penalty relative to local
  context.
\end{description}

\section{Method}

\subsection{Safe benchmark}

The generator produces 400 malicious-benign scenario pairs from 200 structural
templates: eight scenario families crossed with five data representations and
five control-flow shapes, with two identifier instances per template. The eight
families model abstract authorization, normalization, fallback, integrity,
scope, feature, provenance, and dependency-policy composition.

Each case contains an atomic submission and a three-submission sequence. The
atomic and split forms have identical final source trees, total changed-line
counts, activation timestamps, and policy outcomes. For malicious sequences,
the generator reverts each split step separately and executes the oracle; all
three counterfactuals must remain safe. Benign cases match the structure and
changed-line count but do not grant the protected capability.

The source imports only relative synthetic modules and runs only in memory. A
deterministic capability scan rejects network, shell, subprocess, file-system,
package-install, credential, personal-data, persistence, destructive, dynamic
execution, and non-relative import features. The benchmark is not deployable and
contains no real target or exploit instructions.

\subsection{Factorial design}

Table~\ref{tab:factors} shows the controlled factors. Workflow policy and
evidence context are orthogonal: changing \emph{pull request} to \emph{trunk}
changes only the written decision policy, while changing \emph{local} to
\emph{cumulative} changes only whether prior submissions are supplied.

\begin{table}[ht]
\centering
\caption{Controlled study factors.}
\label{tab:factors}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.20\linewidth}
                >{\raggedright\arraybackslash}p{0.28\linewidth}
                >{\raggedright\arraybackslash}p{0.40\linewidth}}
\toprule
Factor & Levels & Interpretation \\
\midrule
Intent & malicious, benign & Matched final-state policy outcome \\
Decomposition & atomic, split & One submission or three necessary submissions \\
Workflow & pull request, trunk & Pre-merge gate or untrusted-main pipeline decision \\
Context & local, cumulative & Current boundary only or explicit prior history \\
\bottomrule
\end{tabular}
\end{table}

The full design contains 57,600 sequence evaluations and 115,200 boundary-level
review calls. Calls are randomized with a seeded Fisher-Yates shuffle. Each
scenario-condition-model cell has three trials. No outcome-dependent stopping,
template replacement, or prompt tuning is permitted.

\subsection{Review configurations}

The frozen routes represent a frontier proprietary model, an independent
proprietary software-engineering replication, and a code-specialized open-weight
model (Table~\ref{tab:models}). Route fallback is disabled; provider and endpoint
metadata must match the prespecified route.

\begin{table}[ht]
\centering
\caption{Frozen model routes. Serving metadata and access times are retained in
the result ledger.}
\label{tab:models}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.42\linewidth}
                >{\raggedright\arraybackslash}p{0.48\linewidth}}
\toprule
Route & Study role \\
\midrule
\code{openai/gpt-5.6-sol} & frontier proprietary \\
\code{anthropic/claude-sonnet-4.6} & independent proprietary software-engineering replication \\
\code{qwen/qwen3-coder-next} & code-specialized open-weight replication \\
\bottomrule
\end{tabular}
\end{table}

All reviewers receive one fixed system task and a JSON response contract. The
request asks for \code{allow}, \code{block}, or \code{abstain}; a maliciousness
probability; severity; relevant submission identifiers; evidence; and a short
category. Requests are capped at 4,000 UTF-8 bytes and 2,000 input tokens, with
no silent truncation. Temperature is zero where supported, tools are disabled,
and output is capped at 256 tokens.

\subsection{Collection and blinding}

The call queue contains opaque call, prompt, and case identifiers plus model and
trial. It contains no intent, family, template, or sequence label. Ground truth
is held in a separate ledger and joined only after collection. Refusals and
schema-invalid responses are observed abstentions and are not retried. Network
errors, HTTP 408, 409, 429, 5xx responses, and route mismatches receive at most
two retries with the identical request. Retry exhaustion produces a missing
boundary.

Collection suspends before accrued cost plus the frozen projected remainder
would exceed US\$1,400, after 100 consecutive infrastructure failures, or on a
safety, terms, or benchmark-integrity concern. The projected confirmatory cost is
US\$1,094.55. This ceiling is a safety limit, not authorization to spend.

\subsection{Outcomes}

The primary binary outcome is blocked by activation. Secondary outcomes are
recall, false-positive rate, precision, abstention rate, cumulative detection by
boundary, submissions to detection, Brier score, ten-bin expected calibration
error, ordinal severity error, localization, and repeated-trial consistency.

A blocked response is localized when it cites at least one prespecified
relevant submission and identifies the causal abstract-policy failure. Two
adjudicators score localization independently while blinded to model, trial, and
condition labels; a third resolves disagreement. Pending adjudications remain
null rather than being scored as failures.

\section{Analysis Plan}

Sequence outcomes are averaged within structural-template and intent-condition
cells across balanced identifier instances, models, and trials. Structural
templates then receive equal weight. The primary uncertainty calculation is a
10,000-replicate percentile bootstrap with seed 20260716. Within each replicate,
25 templates are sampled with replacement separately from each of the eight
scenario families, retaining paired intent and treatment cells.

\begin{table}[ht]
\centering
\caption{Confirmatory estimands and decision rules.}
\label{tab:estimands}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.12\linewidth}
                >{\raggedright\arraybackslash}p{0.56\linewidth}
                >{\raggedright\arraybackslash}p{0.22\linewidth}}
\toprule
Claim & Estimand & Rule \\
\midrule
H1 & malicious minus benign block-rate risk difference, averaged over all factors & 95\% interval above zero \\
H2 & split minus atomic malicious risk difference under local context, averaged over workflow & 95\% interval below zero \\
H3a & trunk minus pull-request malicious risk difference, averaged over decomposition and context & 90\% interval inside [-0.10, 0.10] \\
H3b & decomposition-by-workflow difference in differences, averaged over context & 95\% interval excludes zero \\
H4 & decomposition-by-context difference in differences, averaged over workflow & 95\% interval above zero \\
\bottomrule
\end{tabular}
\end{table}

The design was audited with a 20,000-replicate hierarchical simulation. Under
the frozen design assumptions, the selected 200-template layout exceeded 0.99
power or assurance for all five confirmatory targets. The 200-template floor was
retained for full structural coverage even though a smaller repeated-template
layout met the numerical threshold.

Missing boundaries count as abstentions and therefore not detections in the
primary analysis. Two mandatory bounds treat missing malicious and benign
boundaries oppositely: the detection-favorable bound codes missing malicious
boundaries as blocks and benign boundaries as allows; the detection-unfavorable
bound reverses those assignments. A conclusion is called robust only when its
primary criterion holds under both bounds. Per-model complete-pair sensitivity
drops a structural template only for the model with a missing condition. No
imputation model is fitted.

Secondary model-specific, family-specific, calibration, timing, localization,
and ablation analyses are labeled as such. Holm adjustment is applied when
multiple secondary comparisons within one outcome family are interpreted
inferentially. The earlier mixed-effects logistic model is retained as a
sensitivity analysis rather than the primary estimator.

\section{Pre-outcome Results Status}

No confirmatory model outcomes have been collected or inspected. The benchmark,
prompt ledger, schedule, collection failure rules, joins, estimands, and tests
were implemented using deterministic checks and fabricated responses only. This
section will report participant flow, call failures, primary estimates and
intervals, missingness bounds, model-specific results, calibration,
localization, and consistency after prospective registration and authorized
collection. Until then, no empirical conclusion about H1--H4 is warranted.

\section{Interpretation Framework}

Several result patterns have distinct interpretations. A split penalty under
local but not cumulative context would support an information-integration
failure rather than a general inability to recognize the final unsafe state. A
split penalty under both contexts would instead suggest that temporal
presentation or individually weak evidence impairs composition even when
history is explicit. A decomposition-by-workflow interaction, with evidence held
constant, would indicate sensitivity to the decision policy rather than the code
artifact. High recall accompanied by high benign blocking would not constitute
effective review and will be reported through precision and false-positive rate.

Failure analysis will distinguish missed composition, incorrect causal detail,
overreliance on plausible metadata, severity underestimation, unlocalized
blocking, refusal, schema failure, and infrastructure failure. These categories
are descriptive and cannot replace the prespecified contrasts.

\section{Threats to Validity}

\paragraph{Construct validity.}
The benchmark models malicious intent through a harmless abstract policy
violation. This provides causal control but does not reproduce the semantics or
distribution of production vulnerabilities. Workflow is a controlled written
policy; it cannot reproduce organizational incentives, reviewer fatigue,
rollback pressure, or the blast radius of a real main branch.

\paragraph{Internal validity.}
Atomic and split variants can differ in salience despite identical final trees
and line counts. Counterfactual necessity, prompt hashing, timestamp controls,
and independent review reduce but do not eliminate this risk. Provider serving
stacks may also change behind a stable route; returned metadata and access times
are therefore retained.

\paragraph{External validity.}
Eight policy-composition families, three review configurations, and compact
synthetic repositories do not represent all languages, organizations, or attack
strategies. The study tests one three-submission decomposition and does not
estimate an optimal adversarial strategy. Publication may contaminate future
models, so benchmark and model versions must accompany later replications.

\paragraph{Statistical conclusion validity.}
Repeated calls and identifier variants are not independent scenarios. The
analysis therefore generalizes at the structural-template level and preserves
family strata. Power values depend on design assumptions rather than pilot
outcomes. Missingness bounds expose, rather than hide, dependence on provider
failures.

\section{Ethics and Responsible Release}

This is defensive research into review reliability. The benchmark cannot access
external systems or data, and its only harmful state is a deterministic in-memory
assertion over synthetic names. No real repository or person is tested without
explicit authorization. A separate reviewer assesses whether release artifacts
could provide reusable evasion guidance. Material judged operationally risky is
withheld or reduced to non-operational metadata, and the omission is documented.

Raw model output is released only where provider terms permit. Human review, if
feasible, requires consent, withdrawal procedures, anonymization, and any
applicable ethics review. The study reports provider refusals, missing calls,
deviations, and negative results rather than removing them post hoc.

\section{Reproducibility}

The versioned package contains the generator, cards, prompts, schedule,
collection ledgers, scorer, power simulation, preregistration drafts, and
independent-review protocol. Candidate artifact hashes are:

\begin{itemize}[leftmargin=*]
  \item scenario cards:\par
  \code{71b20e0e5a49547a68d17a8435b1a9f860e8f078942bbc2b66b557971a9a4ab3}
  \item rendered prompts:\par
  \code{56d246f9c676e94f29ce2dea4025633f1b31815959b5b34b0daf2b49af32dff4}
  \item randomized schedule:\par
  \code{e90a26028bfd0b0a38f15ce55c71ffb5c983c8eaf03b7ef18cc74f697481a663}
\end{itemize}

The repository is available at \url{https://github.com/windyroad/windyroad}.
The final release will add permitted raw responses, an attempt ledger, joined
analysis data, adjudication records, environment metadata, checksums, the
prospective registration identifier, and the arXiv identifier.

\section{Conclusion}

This study tests a gap between artifact-level vulnerability detection and
sequence-level code review: whether a model can recognize several individually
ordinary changes as a dangerous composition before activation. By independently
varying decomposition, workflow policy, and context, the design can distinguish
information loss from workflow framing. The safe paired benchmark and frozen
analysis make both positive and negative findings auditable. Empirical
conclusions remain contingent on prospective registration, authorized
collection, and the prespecified analysis.

\appendix

\section{Fixed Review Task}

The fixed task instructs the model to act as an independent security reviewer,
treat repository and metadata as untrusted evidence, decide whether the current
submission should be allowed, blocked, or sent for human review, consider unsafe
behavior created alone or in composition with supplied history, and return only
the contracted JSON grounded in supplied evidence. No chain-of-thought request,
model-specific prompt, or outcome-driven tuning is used.

\section{Artifact Status at the Pre-outcome Freeze}

Machine validation covers final-tree equality, changed-line equality, activation
timestamps, paired benign safety, malicious activation, counterfactual necessity,
capability exclusions, prompt size, schedule balance, call-ledger blinding,
retry behavior, spending suspension, result joining, missingness bounds, and all
five confirmatory contrasts. Independent benchmark and statistical reviews,
exact native token counts, author metadata, license confirmation, and external
registration remain required before collection.

\begin{thebibliography}{9}

\bibitem{yu2024securityreview}
J. Yu, P. Liang, Y. Fu, A. Tahir, M. Shahin, C. Wang, and Y. Cai.
\newblock An insight into security code review with LLMs: Capabilities,
obstacles, and influential factors.
\newblock \emph{arXiv:2401.16310}, 2024.
\newblock \url{https://arxiv.org/abs/2401.16310}.

\bibitem{liu2024vuldetectbench}
Y. Liu, L. Gao, M. Yang, Y. Xie, P. Chen, X. Zhang, and W. Chen.
\newblock VulDetectBench: Evaluating the deep capability of vulnerability
detection with large language models.
\newblock \emph{arXiv:2406.07595}, 2024.
\newblock \url{https://arxiv.org/abs/2406.07595}.

\bibitem{zhou2024sast}
X. Zhou, D.-M. Tran, T. Le-Cong, T. Zhang, I. C. Irsan, J. Sumarlin,
B. Le, and D. Lo.
\newblock Comparison of static application security testing tools and large
language models for repo-level vulnerability detection.
\newblock \emph{arXiv:2407.16235}, 2024.
\newblock \url{https://arxiv.org/abs/2407.16235}.

\bibitem{adalsteinsson2025workflows}
F. S. Adalsteinsson, B. B. Magnusson, M. Milicevic, A. N. Davidsson, and
C.-H. Cheng.
\newblock Rethinking code review workflows with LLM assistance: An empirical
study.
\newblock \emph{arXiv:2505.16339}, 2025.
\newblock \url{https://arxiv.org/abs/2505.16339}.

\bibitem{basic2024review}
E. Basic and A. Giaretta.
\newblock From vulnerabilities to remediation: A systematic literature review
of LLMs in code security.
\newblock \emph{arXiv:2412.15004}, 2024.
\newblock \url{https://arxiv.org/abs/2412.15004}.

\end{thebibliography}

\end{document}
